---
title: Maps with {edgebundle}
author: admin
date: '2020-12-19'
slug: maps-with-edgebundle
categories:
  - blog
  - map
tags:
  - ggplot2
  - tidyverse
  - cartography
  - map
subtitle: 'Mapping US flight networks'
toc: true
summary: ''
authors: []
lastmod: '2020-12-19T13:53:43-06:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  dev = c("CairoPNG"),
  dev.args = list(res = 300, 
                  pointsize = 12)
)

```


## Goal  

Use `{edgebundle}` to map flight patterns from BNA and MSP. 

```{r, pkgs}
# remotes::install_github("schochastics/edgebundle")
library(edgebundle)
library(igraph)
library(ggplot2)
library(ggraph)
library(dplyr)
library(sf)
library(tigris)

set.seed(24601)

my_caption <- c("Liz Roten (@LizRoten) | Data: openflights.org")
```


We also need to use the Python library, datashader. `{edgebundle}` ships with a nice function for installing all the dependencies. 

```{r}
edgebundle:::install_bundle_py()
```


## Data prep

The data we will use is `us_flights`, which is shipped with `{edgebundle}`. `us_flights` is a complex object. 

```{r}
flights <- us_flights # name us_flights
coords <- cbind(V(flights)$longitude, V(flights)$latitude) # extract coordinates

# create vertex sequence
verts <- data.frame(x = V(flights)$longitude, y = V(flights)$latitude) 
```

### Supporting data

To make our output a little more aesthetically pleasing, we will go ahead and transform the data to use Albers Equal Area Conic.  

```{r}
states <- tigris::states(cb = TRUE, progress_bar = FALSE) %>% 
  filter(STUSPS %in% state.abb,
         !NAME %in% c("Alaska",
                      "Hawaii")) %>% 
  sf::st_transform(crs = " +proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs")
```

```{r}
coords_full <- cbind(V(flights)$longitude, V(flights)$latitude,  V(flights)$name) # extract coordinates


coords_sf <- st_as_sf(x = as.data.frame(coords_full), coords = c("V1", "V2"), crs = 4326) %>% 
  sf::st_transform(crs = " +proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs")

```

## Edge bundles  
Create edge bundles

```{r}
force_bundle <- edge_bundle_force(flights, xy = coords, compatibility_threshold = 0.6) 

force_bundle_sf <- force_bundle %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% 
  sf::st_transform(crs = " +proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs") %>% 
  rowwise() %>% 
  mutate(x_coord = st_coordinates(geometry)[[1]],
         y_coord = st_coordinates(geometry)[[2]])

```

```{r, cache=TRUE}
hammer_bundle <- edgebundle::edge_bundle_hammer(object = flights, 
                                                xy = coords, 
                                                bw = 0.7,
                                                decay = 0.5)

hammer_bundle_sf <- hammer_bundle %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% 
  sf::st_transform(crs = " +proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs") %>% 
  rowwise() %>% 
  mutate(x_coord = st_coordinates(geometry)[[1]],
         y_coord = st_coordinates(geometry)[[2]])
```



## Map

```{r}
my_theme <- theme_void() +
  theme(
    plot.background = element_rect(fill = "black",
                                   linetype = 0,
                                   color = NA),
    panel.background = element_rect(
      fill = "black",
      linetype = 0,
      color = NA
    ),
    plot.title = element_text(family = "Agency FB",
                              color = "white",
                              size = 20),
    plot.caption = element_text(
      family = "LeelawadeeUI-Semilight",
      color = "white"
    ),
    plot.subtitle = element_text(
      color = "white"
    ),
    plot.margin = unit(c(.5, .5, .2, .5), "cm"),
  )

line_color <- "#e79e4f"
```


```{r}
base_plot <- geom_sf(data = states,
          color = "white",
          fill = NA,
          lwd = 0.1) 
```


```{r "force_bundle_method"}
ggplot() +
  base_plot +
  geom_path(data = force_bundle_sf,
            aes(x = x_coord,
                y = y_coord,
                group = group),
            color = line_color,
            size = 0.5,
            alpha = 0.2) +
  geom_path(data = force_bundle_sf,
            aes(x = x_coord,
                y = y_coord,
                group = group),
            color = "white",
            size = 0.005,
            alpha = 0.1) +
  geom_sf(data = coords_sf,
          color = line_color,
          size = 0.25) +
    geom_sf(data = coords_sf,
          color = "white",
          size = 0.25,
          alpha = 0.1) +
  labs(title = "US Flight Network",
       subtitle = "Force Bundle Method",
       caption = my_caption) +
  my_theme 

```

```{r "hammer_bundle_method"}
ggplot() +
  base_plot +
    geom_path(data = hammer_bundle_sf,
            aes(x = x_coord,
                y = y_coord,
                group = group),
            color = line_color,
            size = 0.5,
            alpha = 0.2) +
  geom_path(data = hammer_bundle_sf,
            aes(x = x_coord,
                y = y_coord,
                group = group),
            color = "white",
            size = 0.005,
            alpha = 0.1) +
  geom_sf(data = coords_sf,
          color = line_color,
          size = 0.25) +
    geom_sf(data = coords_sf,
          color = "white",
          size = 0.25,
          alpha = 0.1) +
  labs(title = "US Flight Network",
       subtitle = "Hammer Bundle Method",
       caption = my_caption) +
  my_theme 
```

## Finese hammer bundle method  


```{r, "hammer_bundle_method_adj"}
hammer_bundle_adj <- edgebundle::edge_bundle_hammer(object = flights, 
                                                xy = coords, 
                                                bw = 0.2,
                                                decay = 0.5)

hammer_bundle_adj_sf <- hammer_bundle_adj %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% 
  sf::st_transform(crs = " +proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs") %>% 
  rowwise() %>% 
  mutate(x_coord = st_coordinates(geometry)[[1]],
         y_coord = st_coordinates(geometry)[[2]])

ggplot() +
  base_plot +
    geom_path(data = hammer_bundle_adj_sf,
            aes(x = x_coord,
                y = y_coord,
                group = group),
            color = line_color,
            size = 0.5,
            alpha = 0.2) +
  geom_path(data = hammer_bundle_adj_sf,
            aes(x = x_coord,
                y = y_coord,
                group = group),
            color = "white",
            size = 0.005,
            alpha = 0.1) +
  geom_sf(data = coords_sf,
          color = line_color,
          size = 0.25) +
    geom_sf(data = coords_sf,
          color = "white",
          size = 0.25,
          alpha = 0.1) +
  labs(title = "US Flight Network",
       subtitle = "Hammer Bundle Method (adj)",
       caption = my_caption) +
  my_theme 
```

